---
title: "Summarising EDA Features"
output:
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidyverse)
library(purrr)
library(stringr)
library(furrr)
library(lubridate)
library(broom)
library(knitr)
library(readxl, quietly = TRUE)
```

In this notebook we summarise and visualise the EDA features we extracted.

```{r}
eda_df <- list.files("/Users/mac/GDriveDocs/Quantitative Psychology & Statistics Lab/Emote/EDA_Analysis/outputs",
                     pattern = ".rds",
                     recursive = TRUE,
                     full.names = TRUE) %>%
  future_map_dfr(.f = function(x) {
    
    df <- readRDS(x)
    df$ID <- str_extract(x, "[[:digit:]]+")
    
    df
  })

events = read_excel("/Users/mac/GDriveDocs/Quantitative Psychology & Statistics Lab/Emote/DataOutputs/UpdatedSurveys_180713.xlsx") %>%
  group_by(ID) %>%
  arrange(ID, Event) %>%
  ungroup() %>%
  filter(!is.na(ID)) %>%
  mutate(Event = ymd_hms(Event)) %>%
  mutate(start = Event - 60 * 30) %>%
  mutate(intervals = interval(Event, start))
  

remove_overlaps <- function(intervals){
  
  overlaps = rep(FALSE, length(intervals))
  
  for(l in 1:(length(intervals)-1)){
    if(!(is.na(intervals[l]) & is.na(intervals[l+1]))){
      if(int_overlaps(intervals[l], intervals[l+1])){
        overlaps[l] = TRUE
      }
    }
  }
  
  return(overlaps)
}

events <- events %>%
  mutate(ID = as.factor(ID)) %>%
  group_by(ID) %>%
  mutate(overlapped = remove_overlaps(intervals)) %>%
  select(-intervals) %>%
  filter(overlapped ==  FALSE) %>%
  select(-overlapped) %>%
  ungroup()

eda_df <- events %>%
  select(ID, EmotionalEatingEpisodes, Event, Stress) %>%
  mutate(ID = as.character(ID)) %>%
  left_join(eda_df, by = c("ID", "Event"))
```

# Accelerometer Features

## Step Count

`Step count` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Step count`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$`Step count`))) %>%
  kable(caption = "\"Step Count\" Summary During Episodes and Controls")
```

`Step count` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Step count`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y == "Episode")%>% #filter high stress
  do(tidy(summary(.$`Step count`))) %>%
  kable(caption = "\"Step Count\" Summary During Episodes")
```

`Step count` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Step count`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$`Step count`))) %>%
  kable(caption = "\"Step Count\" Summary During Controls")
```


Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Step count`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=`Step count`)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Visualising Step Count") +
  theme_minimal() +
  coord_flip()
```

## Mean step time during observations

`Mean step time during observation` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Mean step time during movement`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$`Mean step time during movement`))) %>%
  kable(caption = "\"Mean step time during observation\" Summary During Episodes and Controls")
```

`Mean step time during observation` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Mean step time during movement`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Episode") & Y == "Episode")%>% #filter high stress
  do(tidy(summary(.$`Mean step time during movement`))) %>%
  kable(caption = "\"Mean step time during observation\" Summary During Episodes")
```

`Mean step time during observation` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Mean step time during movement`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$`Mean step time during movement`))) %>%
  kable(caption = "\"Mean step time during observation\" Summary During Controls")
```


Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Mean step time during movement`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=`Mean step time during movement`)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Visualising Mean step time during observation",
       caption = "Time in seconds") +
  theme_minimal() +
  coord_flip()
```

## Percent Stillness during observations

`Percent Stillness` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Percent stillness`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$`Percent stillness`))) %>%
  kable(caption = "\"Percent stillness during observation\" Summary During Episodes and Controls")
```

`Percent stillness during observation` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Percent stillness`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Episode") & Y == "Episode")%>% #filter high stress
  do(tidy(summary(.$`Percent stillness`))) %>%
  kable(caption = "\"Percent stillness during observation\" Summary for Episodes")
```

`Percent stillness during observation` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Percent stillness`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$`Percent stillness`))) %>%
  kable(caption = "\"Percent stillness during observation\" Summary for Controls")
```


Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         `Percent stillness`,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=`Percent stillness`)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Visualising Percent stillness during observations") +
  theme_minimal() +
  coord_flip()
```

# Temperature

`Temperature` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         Temp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$Temp))) %>%
  kable(caption = "\"Temperature during observation\" Summary During Episodes and Controls")
```

`Temperature during observation` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         Temp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Episode") & Y == "Episode") %>% #filter high stress
  do(tidy(summary(.$Temp))) %>%
  kable(caption = "\"Temperature during observation\" Summary for Episodes")
```

`Temperature during observation` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         Temp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$Temp))) %>%
  kable(caption = "\"Temperature during observation\" Summary for Controls")
```

Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         Temp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=Temp)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Visualising Temperature during observations") +
  theme_minimal() +
  coord_flip()
```

Lots of outliers =D

# Motion

## Aggregate motion during observations

This is `sqrt(X^2+Y^2+Z^2)`

`Motion` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         motion,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$motion))) %>%
  kable(caption = "\"3 dimensional motion during observation\" Summary During Episodes and Controls")
```

`Motion during observation` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         motion,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y == "Episode") %>% #filter high stress
  do(tidy(summary(.$motion))) %>%
  kable(caption = "\"3 dimensional motion during observation\" Summary for Episodes")
```

`Motion during observation` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         motion,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$motion))) %>%
  kable(caption = "\"3 dimensional motion during observation\" Summary for Controls")
```

Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         motion,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=motion)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Visualising 3 Dimensional motion during observations") +
  theme_minimal() +
  coord_flip()
```

Even more outliers! =D

# EDA Peaks

## Number of Peaks by Participant

```{r}
eda_df %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         peaks,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control"),
         peaks = as.numeric(peaks)) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  group_by(ID) %>%
  summarise(peaks2 = sum(peaks, na.rm = TRUE)) %>%
  ggplot(aes(x=ID,y=peaks2)) +
    geom_col()+
    theme_minimal() +
    labs(title = "Number of EDA Peaks per Participant", y = "Number of Peaks")

```

## Number of Peaks by Condition

```{r}
eda_df %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         peaks,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control"),
         peaks = as.numeric(peaks)) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  group_by(ID, Y) %>%
  summarise(peaks2 = sum(peaks, na.rm = TRUE)) %>%
  ggplot(aes(x=ID,y=peaks2)) +
    geom_col(aes(fill = Y), position = "dodge") +
    theme_minimal() +
    labs(title = "Number of EDA Peaks per Participant & Condition", y = "Number of Peaks")

```

# EDA Values

## Amplitude

`Peak Amplitude` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         amp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$amp))) %>%
  kable(caption = "\"EDA Peak Amplitude\" Summary During Episodes and Controls")
```

Peak amplitude summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         amp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y == "Episode")%>% #filter high stress
  do(tidy(summary(.$amp))) %>%
  kable(caption = "\"Peak Amplitude\" Summary During Episodes")
```

`Peak Amplitude` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         amp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$amp))) %>%
  kable(caption = "\"Peak Amplitude\" Summary During Episodes")
```


Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         amp,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=amp)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Peak Amplitudes during observations") +
  theme_minimal() +
  coord_flip()
```

## Max Derivative

This is defined as max derivative within 1 second of apex of SCR

`Max Deriv` summaries for both episodes and controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         max_deriv,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  do(tidy(summary(.$max_deriv))) %>%
  kable(caption = "\"Max Derivative\" Summary During Episodes and Controls")
```

`Max Derivative` summaries for episodes:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         max_deriv,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y == "Episode")%>% #filter high stress
  do(tidy(summary(.$max_deriv))) %>%
  kable(caption = "\"Max Derivative\" Summary During Episodes")
```

`Max Derivative` summaries for controls:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         max_deriv,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control") & Y != "Episode")%>% #filter high stress
  do(tidy(summary(.$max_deriv))) %>%
  kable(caption = "\"Max Derivative\" Summary During Episodes")
```


Visualised:
```{r}
eda_df %>%
  group_by(ID) %>%
  select(ID, 
         Event, 
         EmotionalEatingEpisodes,
         max_deriv,
         Stress) %>%
  distinct() %>%
  mutate(Y = ifelse(EmotionalEatingEpisodes == 1, "Episode", "Control")) %>%
  filter(!(Stress > 4 & Y == "Control")) %>%
  ggplot(aes(x=ID,y=max_deriv)) +
  geom_boxplot(aes(fill = Y)) +
  labs(title = "Max Derivative Features during observations") +
  theme_minimal() +
  coord_flip()
```

