---
title: "EDA Feature Extraction"
output: html_notebook
---

In this notebook we demonstrate how to collect and organise the EDA features into usable observations

```{r, session settings, echo=FALSE}
TZ = "America/New_York"
Sys.setenv(TZ="America/New_York")
library(tidyverse, quietly = TRUE) #for tidy R programming
library(data.table, quietly = TRUE) #for fast csv reading
library(lubridate, quietly = TRUE) #for datetime manipulation
library(stringr, quietly = TRUE)
library(readr, quietly = TRUE)
library(purrr, quietly = TRUE)
```

First get the survey data:
```{r}
events <- read_csv("/Users/mac/GDriveDocs/Quantitative Psychology & Statistics Lab/Emote/DataOutputs/UpdatedSurveys_180713.csv")

events$Event <- ymd_hms(events$Event)
head(events)
```

Then, for one user, read in their accelerometer features:
```{r}
p200 <- "/Users/mac/GDriveDocs/Quantitative Psychology & Statistics Lab/Emote/EDA_Analysis/outputs/200"

acc_features <- list.files(p200, full.names = TRUE) %>%
  grep(pattern = "features", value = TRUE) %>%
  read_csv()

head(acc_features)
```

And the list of data frames of their EDA peaks
```{r}
read_and_name <- function(f){
  
  csv <- read_csv(f)
  event <- ymd_hms(str_replace(basename(f), "\\.csv", ""))
  csv$Event = event
  
  csv
}

eda_peaks <- list.files(p200, full.names = TRUE) %>%
  grep(pattern = "features", value = TRUE, invert = TRUE) %>%
  map_df(., read_and_name)

eda_peaks <- dplyr::arrange(eda_peaks, Event)
head(eda_peaks)
```

To make it easy, we will join the peak data and accelerometer data:
```{r}
full_eda <- left_join(acc_features, eda_peaks, by = "Event")

str(full_eda)
```

We can write this file out to disk as an .rds

Now, we can loop through the participants to apply this strategy

```{r}
participants <- list.dirs("/Users/mac/GDriveDocs/Quantitative Psychology & Statistics Lab/Emote/EDA_Analysis/outputs", full.names = TRUE) %>%
  grep(pattern = "[[:digit:]]$", value = TRUE)

for(x in participants) {
  
  acc_features <- list.files(x, full.names = TRUE) %>%
    grep(pattern = "features", value = TRUE) %>%
    read_csv()
  
  eda_peaks <- list.files(x, full.names = TRUE) %>%
    grep(pattern = "features", value = TRUE, invert = TRUE) %>%
    grep(pattern = "csv", value = TRUE) %>%
    map_df(., read_and_name)
  
  
  full_eda <- left_join(acc_features, eda_peaks, by = "Event")
  
  write_rds(full_eda, paste0(x, "/full_out.rds"))
}
```

