---
title: "FrequencyDomainSummary3"
output: html_notebook
---

# Identify the outliers in frequency domain and observe individually

The following rows have at least one value where there is a positive outlier (1.5 $\times$ the interquartile range).

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(when, ID, Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(Avg_niHR:LFHF_6), any_vars(. > IQR(., na.rm = T)*1.5))
```

Moving forward, we filter control outliers out, and truncate episode outliers, filtering by highest value possible according to the literature.

HF max ~ 3630

LF max ~ 1009

LFHF max ~ 11.6


The following table shows all of the available data (no row-wise completeness required) when we apply the above filters.

```{r}
HF_max = 3630
LF_max = 1009
LFHF_max = 11.6
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%#if is episode and outlier, assign to max val
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

Split by condition we have, for controls:
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%#if is episode and outlier, assign to max val
  filter(Y == "Control")%>%
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

And for episodes:
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%#if is episode and outlier, assign to max val
  filter(Y != "Control")%>%
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

# Visualisations After Filtering & No Row-Wise Completion

All plots shown for full data set and split by condition.

## Heart Rate
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%#if is episode and outlier, assign to max val
  gather("Variable", "Value", 2:5)%>%
  ggplot(aes(Value))+
  geom_density(aes(colour=Variable), alpha=0.5)+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "Heart Rate Variables")+
  scale_fill_viridis_d()
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%#if is episode and outlier, assign to max val
  gather("Variable", "Value", 2:5)%>%
  ggplot(aes(Value))+
  geom_density(aes(colour=Variable), alpha=0.5)+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "Heart Rate Variables")+
  scale_fill_viridis_d()+
  facet_grid(Y~.)
```

## High frequency

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, HF_1:HF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  geom_density(aes(colour=Variable), alpha=0.5)+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "High Frequency Variables")+
  scale_fill_viridis_d()
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, HF_1:HF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  geom_density(aes(colour=Variable), alpha=0.5)+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "High Frequency Variables")+
  scale_fill_viridis_d()+
  facet_grid(Y~.)
```

## Low frequency
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, LF_1:LF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  geom_density(aes(colour=Variable), alpha=0.5)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "Low Frequency Variables")+
  scale_fill_viridis_d()
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, LF_1:LF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  geom_density(aes(colour=Variable), alpha=0.5)+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "Low Frequency Variables")+
  scale_fill_viridis_d()+
  facet_grid(Y~.)
```

## LFHF

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, LFHF_1:LFHF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  geom_density(aes(colour=Variable))+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "LF/HF Variables")+
  scale_fill_viridis_d()
```
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  select(Y, LFHF_1:LFHF_6)%>%
  #filter(complete.cases(.))%>%
  # group_by(ID)%>%
  # mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  # ungroup()%>%
  #select(Y, everything())%>%
  gather("Variable", "Value", 2:7)%>%
  ggplot(aes(Value))+
  #geom_histogram(aes(fill=Variable), alpha=0.5, bins = 100)+
  geom_density(aes(colour=Variable))+
  theme_minimal()+
  labs(title = "Histogram of All Variables", subtitle = "LF/HF Variables")+
  scale_fill_viridis_d()+
  facet_grid(Y~.)
```

# Final Amount of Data

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  count()
```

# Between Groups Tests

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(Y, Avg_niHR:LFHF_6)%>%
  filter_at(.vars = vars(HF_1:HF_6), any_vars(!(Y == "Control" & . > HF_max)))%>%#if is control and outlier, remove
  filter_at(.vars = vars(LF_1:LF_6), any_vars(!(Y == "Control" & . > LF_max)))%>% #if is control and outlier, remove
  filter_at(.vars = vars(LFHF_1:LFHF_6), any_vars(!(Y == "Control" & . > LFHF_max)))%>% #if is control and outlier, remove
  mutate_at(.vars = vars(HF_1:HF_6), funs(ifelse(. > HF_max, HF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LF_1:LF_6), funs(ifelse(. > LF_max, LF_max, .)))%>%#if is episode and outlier, assign to max val
  mutate_at(.vars = vars(LFHF_1:LFHF_6), funs(ifelse(. > LFHF_max, LFHF_max, .)))%>%
  summarise_at(funs(twotPermutation(.[Y == "Control" & !is.na(.)],
                                    .[Y != "Control" & !is.na(.)], 
                                    plotit = F)),
               .vars = vars(Avg_niHR:LFHF_6))%>%
  gather("Variable", "P.Value")%>%
  mutate(Significant = ifelse(P.Value < 0.05, "Significant", "Not Significant"))
```

