---
title: "FrequencyDomainSummary2"
output: html_notebook
---

# Frequency Domain Summary 2
How many cases are there to use?
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  select(ID, Y, Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  count()
```

```{r}
#save participants who qualify with both episode and control
pp = c(201, 202, 207, 214, 215, 217, 218, 219, 226, 228)
```

# Within Participant Comparisons

Here we compare within-participant if episodes have different frequency-scale variables from controls, for each participant.

A plot for clarity
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  ungroup()%>%
  select(Y, everything())%>%
  gather("Variable", "Value", 3:6)%>%
  ggplot(aes(x=Variable, y=Value))+
  geom_boxplot(aes(fill=Y))+
  theme_minimal()+
  labs(y = "Z Score", title = "Boxplot of All Variables By Condition", subtitle = "Heart Rate Variables")+
  facet_grid(ID~.)
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  ungroup()%>%
  select(Y, everything())%>%
  gather("Variable", "Value", 7:12)%>%
  ggplot(aes(x=Variable, y=Value))+
  geom_boxplot(aes(fill=Y))+
  theme_minimal()+
  labs(y = "Z Score", title = "Boxplot of All Variables By Condition", subtitle = "High Frequency Domain")+
  facet_grid(ID~.)
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  ungroup()%>%
  select(Y, everything())%>%
  gather("Variable", "Value", 13:18)%>%
  ggplot(aes(x=Variable, y=Value))+
  geom_boxplot(aes(fill=Y))+
  theme_minimal()+
  labs(y = "Z Score", title = "Boxplot of All Variables By Condition", subtitle = "Low Frequency Domain")+
  facet_grid(ID~.)
```

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(scale))%>% #z score transform
  ungroup()%>%
  select(Y, everything())%>%
  gather("Variable", "Value", 6:9)%>%
  ggplot(aes(x=Variable, y=Value))+
  geom_boxplot(aes(fill=Y))+
  theme_minimal()+
  labs(y = "Z Score", title = "Boxplot of All Variables By Condition", subtitle = "LFHF")+
  facet_grid(ID~.)
```

Single lines indicate that there's only one reading for that participant's variable. This is in a lot of cases, unfortunately, as you can see from the previous table. I'd only feel comfortable running the permutation T with 207. But we can still attempt the t-tests regardless:

```{r}
simple_scale = function(vec){
  vec%>%
    scale()%>%
    as.numeric()%>%
    return()
}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  #filter(ID == 207)%>%
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(simple_scale))%>% #z score transform
  select(Y:LFHF_6)%>%
  summarise_at(funs(twotPermutation(.[Y == "Control"],
                                    .[Y != "Control"], 
                                    plotit = F)),
               .vars = vars(Avg_niHR:LFHF_6))
```

The above table shows the $p$ value for each Episode vs Control hypothesis test that the variable is significantly different. Each row is a participant, and each column is a variable to be tested for condition. None of the tests were significant within participants. Using the full dataset we test it again:

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  #filter(ID == 207)%>%
  filter(ID %in% pp)%>%
  select(ID, Y, Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  #group_by(ID)%>%
  mutate_at(.vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6), funs(simple_scale))%>% #z score transform
  select(Y:LFHF_6)%>%
  #ungroup()%>%
  summarise_at(funs(twotPermutation(.[Y == "Control"],
                                    .[Y != "Control"], 
                                    plotit = F)),
               .vars = vars(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6))#%>%
  #filter_at(vars(Avg_niHR:LFHF_6), any_vars(. < 0.05))
```

None are significant. 

We can still get a quick summary table of all the features though. In total:

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  #filter(ID == 207)%>%
  filter(ID %in% pp)%>%
  select(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

For episodes:
```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(Y == "Episode")%>%
  filter(ID %in% pp)%>%
  select(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

For controls:

```{r}
features_output%>%
  filter(!(Stress > 4 & Y == "Control"))%>% #filter high stress
  filter(Y != "Episode")%>%
  filter(ID %in% pp)%>%
  select(Avg_niHR:Avg_HR, HF_6, LF_6, LFHF_6)%>%
  filter(complete.cases(.))%>%
  psych::describe(quant=c(.25,.75))%>%
  as_tibble()%>%
  rownames_to_column()%>%
  rename(Variable = rowname)%>%
  select(-c(vars, trimmed, mad))
```

