---
title: "Analysing Heart Rate"
output: html_notebook
---

We quickly examine heart rate in this dataset. Note that the difference between instantaneous HR (niHR) and regular HR (HR) is that the niHR is defined by the current and surrounding IBI's so it is a reflection of the last 2-3 cardiac beats. The true heart rate is over a longer period of time (in our experiment, 30 minutes).
```{r, message=FALSE, warning=FALSE, include=FALSE}
DURATION = 60*30
#THRESHOLD = 30

ibi_outputs = data.frame(threshold = NULL, duration = NULL, index = NULL, ibi_mean = NULL, ibi_min = NULL, ibi_max = NULL)
features_output = data.frame()

min_windows = c(2)
thresholds = c(250)

for(th in 1:length(thresholds)){
  THRESHOLD = thresholds[th]
  
  for(w in 1:length(min_windows)){
  
    MIN_WINDOWS = min_windows[w]
    for(i in 1:nrow(surveys)){
    
      example_survey = surveys[i,]
      
      rhrv = ibi_data%>%
        filter(ID == example_survey$ID[1])%>%
        filter(Beat <= example_survey$Event[1] & Beat >= example_survey$Event[1] - DURATION)
      
      if(nrow(rhrv) < 60){
          
        ibi_outputs = data.frame(threshold = THRESHOLD, winds = MIN_WINDOWS, index = i, ibi_mean = NA, ibi_min = NA, ibi_max = NA)%>%
          rbind(ibi_outputs)
        next
      
      }else{
        
        rhrv2 = MakeHRVObject(rhrv)
        
        ibi_outputs = data.frame(threshold = THRESHOLD, 
                                 winds = MIN_WINDOWS, index = i, 
                                 ibi_mean = mean(rhrv2$Beat$RR),
                                 ibi_min = min(rhrv2$Beat$RR), 
                                 ibi_max = max(rhrv2$Beat$RR))%>%
          rbind(ibi_outputs)
        
        if(nrow(rhrv2$Beat) < 60){
    
          time_analysis = data.frame(threshold = THRESHOLD, 
                                     winds = MIN_WINDOWS, 
                                     index = i,
                                     Avg_niHR = NA,
                                     Start_niHR = NA,
                                     End_niHR = NA,
                                     Avg_HR = NA,
                                     SDNN = NA,
                                     SDANN = NA,
                                     SDNNIDX = NA,
                                     pNN50 = NA,
                                     SDSD = NA,
                                     rMSSD = NA,
                                     IRRR = NA,
                                     MADRR = NA,
                                     TINN = NA,
                                     HRVi = NA)
          
        }else{
          time_analysis = cbind(threshold = THRESHOLD, winds = MIN_WINDOWS, index = i,
                                ExtractTimeFeatures(rhrv2, time_window_size = 300,
                                                    minNumWindows = MIN_WINDOWS,
                                                    windowBeatMin = THRESHOLD))
          
        }
        
        features_output = rbind(features_output, cbind(example_survey, time_analysis))
      }
    }
  }
}


```

Again, the dataset looks like this:
```{r}
features_output%>%
  filter(threshold == 250 & winds == 2)%>%
  select(-winds, - threshold, -index)%>%
  filter(complete.cases(.))%>%
  summary()
```

We filter out the stress outlier and examine the statistics of niHR and HR.

# Investigating Heart Rate by Condition
```{r}
features_output%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  filter(!(abs(Stress - median(Stress)) > 2*sd(Stress)))%>% #pull stress outlier
  select(Y,Avg_niHR:Avg_HR)%>%
  gather("Variable", "Value", 2:5)%>%
  ggplot(aes(x=Variable, y=Value))+
  geom_boxplot(aes(fill=Y))+
  theme_minimal()+
  labs(title = "Heart Rate Statistics By Condition")
```

These are the exact mean values:

```{r}
features_output%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  filter(!(abs(Stress - median(Stress)) > 2*sd(Stress)))%>% #pull stress outlier
  select(Y,Avg_niHR:Avg_HR)%>%
  summarise_at(funs(mean(.)),
               .vars = vars(Avg_niHR:Avg_HR))
```

Are any of these values statistically different? $p$-values shown below:
```{r}
features_output%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  filter(!(abs(Stress - median(Stress)) > 2*sd(Stress)))%>% #pull stress outlier
  ungroup()%>%
  select(Y,Avg_niHR:Avg_HR)%>%
  summarise_at(funs(t.test(x = .[Y == "Control"],
                           y = .[Y != "Control"])$p.value),
               .vars = vars(Avg_niHR:Avg_HR))
```

And using the permutation test:
```{r}
features_output%>%
  filter(complete.cases(.))%>%
  group_by(Y)%>%
  filter(!(abs(Stress - median(Stress)) > 2*sd(Stress)))%>% #pull stress outlier
  ungroup()%>%
  select(Y,Avg_niHR:Avg_HR)%>%
  summarise_at(funs(twotPermutation(.[Y == "Control"],
                                    .[Y != "Control"], 
                                    plotit = F)),
               .vars = vars(Avg_niHR:Avg_HR))
```

Doesn't look like heart rate is statistically different between conditions. Oh well.

