---
title: "Participant207"
output: html_notebook
---

# A Closer Look at Participant 207's Data

## IBI

The surveys that 207 had sufficient data for either time or frequency domain:
```{r}
#index of 207's surveys
idx_time = features_output%>%
  filter(ID == 207)%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  select(index)%>%
  .$index

idx_freq = features_output%>%
  filter(ID == 207)%>%
  filter_at(vars(Avg_niHR:LFHF_6), all_vars(complete.cases(.)))%>%
  select(index)%>%
  .$index

features_output%>%
  filter(index %in% idx_time | index %in% idx_freq)
```

26 surveys, 9 of which were episodes. What is the IBI?

```{r}
ibi_outputs%>%
  left_join(features_output, by="index")%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  select(ID, ibi_mean:ibi_max)%>%
  mutate(ID = ifelse(ID == 207, "207", "Other"))%>%
  gather("key", "value", 2:4)%>%
  ggplot(aes(y=value, x=key))+
  geom_boxplot(aes(fill = ID), alpha = 0.5, outlier.colour = "red", outlier.size = 5)+
  #geom_dotplot(binaxis = "y", stackdir = "center", alpha = 0.5)+
  theme_minimal()+
  labs(title = "IBI Values for All Completed Surveys", 
       subtitle = "Compares Participant 207 vs. All Other Participants")
```

207 seems to have lowered IBI values for all of their complete data.

## BMI

```{r}
library(foreign)
qual = read.spss("/Volumes/psychology/Shared/WEACT/EMOTE/Data/EmoteMASTERdatabase.sav", to.data.frame = T)

qual%>%
  select(ID, BMIA1)%>%
  left_join(features_output, by="ID")%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  filter(complete.cases(.))%>%
  select(ID, BMIA1)%>%
  unique()%>%
  mutate(ID = as.factor(ID))%>%
  ggplot(aes(x=ID, y=BMIA1))+
  geom_col()+
  theme_minimal()+
  labs(title = "Participants' BMI at A1")
```

Participant 207 doesn't have abnormally low or high BMI.

## Survey Time

```{r}
require(lubridate)
require(viridis)
ibi_outputs%>%
  left_join(features_output, by="index")%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  select(ID, when)%>%
  mutate(ID = ifelse(ID == 207, "207", "Other"),
         day = wday(when, label = T))%>%
  ggplot(aes(x = day, fill=ID))+
  geom_bar(aes(group=ID), stat="count", position = "stack")+
  theme_minimal()+
  labs(title="Number of Completed Surveys By Day of Week and Participant")+
  scale_fill_viridis(discrete=TRUE)

ibi_outputs%>%
  left_join(features_output, by="index")%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  select(ID, when)%>%
  mutate(ID = ifelse(ID == 207, "207", "Other"),
         hour = as.factor(hour(when)))%>%
  ggplot(aes(x = hour, fill=ID))+
  geom_bar(aes(group=ID), stat="count", position = "stack")+
  theme_minimal()+
  labs(title="Number of Completed Surveys By Hour of Day and Participant")+
  scale_fill_viridis(discrete=TRUE)
```

Participant 207 doesn't seem to be completing their surveys differently to anybody else. They have a large proportion of Monday surveys but not extraneously so.

# Time Domain Features

```{r}
ibi_outputs%>%
  left_join(features_output, by="index")%>%
  filter_at(vars(Y:HRVi), all_vars(complete.cases(.)))%>%
  mutate(ID = ifelse(ID == 207, "207", "Other"))%>%
  select(ID, SDNN:HRVi)%>%
  mutate_at(.vars = vars(SDNN:HRVi), funs(scale))%>%
  gather("variable", "value", SDNN:HRVi)%>%
  ggplot(aes(value, color = variable))+
  geom_density(alpha = 0.5)+
  facet_grid(rows=vars(ID))+
  theme_minimal()+
  scale_color_viridis(discrete = TRUE)+
  labs(title = "Density Plots of Z Score Variables for All Surveys",
       subtitle = "Compares Participant 207 Against All Others")
```

There doesnt seem to be much difference here either.

## Heart Rate

```{r}
ibi_outputs%>%
  left_join(features_output, by="index")%>%
  filter_at(vars(Avg_niHR:Avg_HR), all_vars(complete.cases(.)))%>%
  select(ID, Avg_niHR:Avg_HR)%>%
  mutate(ID = ifelse(ID == 207, "207", "Other"))%>%
  gather("key", "value", 2:5)%>%
  ggplot(aes(y=value, x=key))+
  geom_boxplot(aes(fill = ID), alpha = 0.5, outlier.colour = "red", outlier.size = 5)+
  #geom_dotplot(binaxis = "y", stackdir = "center", alpha = 0.5)+
  theme_minimal()+
  labs(title = "Heart Rate for All Completed Surveys", 
       subtitle = "Compares Participant 207 vs. All Other Participants")
```

Participant seems to have a more elevated heart rate as compared to the rest of the data.